---
title: "Principle Component Aanalysis"
output: html_document
---
## Data
The data you will be using comes from the Assistments online intelligent tutoring system (https://www.assistments.org/). It describes students working through online math problems. Each student has the following data associated with them:

- id
- prior_prob_count: How many problems a student has answered in the system prior to this session
- prior_percent_correct: The percentage of problems a student has answered correctly prior to this session
- problems_attempted: The number of problems the student has attempted in the current session
- mean_correct: The average number of correct answers a student made on their first attempt at problems in the current session
- mean_hint: The average number of hints a student asked for in the current session
- mean_attempt: The average number of attempts a student took to answer a problem in the current session
- mean_confidence: The average confidence each student has in their ability to answer the problems in the current session

## Start by uploading the data
```{r}
D1 <- read.csv("Assistments-confidence.csv", header = TRUE)

```

## Create a correlation matrix of the relationships between the variables, including correlation coefficients for each pair of variables/features.

```{r}
#You can install the corrplot package to plot some pretty correlation matrices (sometimes called correlograms)

library(ggplot2)
library(GGally)

ggpairs(D1, 2:8, progress = FALSE) #ggpairs() draws a correlation plot between all the columns you identify by number (second option, you don't need the first column as it is the student ID) and progress = FALSE stops a progress bar appearing as it renders your plot

ggcorr(D1[,-1], method = c("everything", "pearson")) #ggcorr() doesn't have an explicit option to choose variables so we need to use matrix notation to drop the id variable. We then need to choose a "method" which determines how to treat missing values (here we choose to keep everything, and then which kind of correlation calculation to use, here we are using Pearson correlation, the other options are "kendall" or "spearman")

#Study your correlogram images and save them, you will need them later. Take note of what is strongly related to the outcome variable of interest, mean_correct. 
```

## Create a new data frame with the mean_correct variable removed, we want to keep that variable intact. The other variables will be included in our PCA.

```{r}
# Remove id and mean_correct variable
D2 <- D1[, -c(1, 5)]

```

## Now run the PCA on the new data frame

```{r}
pca <- prcomp(D2, scale. = TRUE)
```

## Although princomp does not generate the eigenvalues directly for us, we can print a list of the standard deviation of the variance accounted for by each component.

```{r}
pca$sdev

#To convert this into variance accounted for we can square it, these numbers are proportional to the eigenvalue

pca$sdev^2

#A summary of our pca will give us the proportion of variance accounted for by each component

summary(pca)

#We can look at this to get an idea of which components we should keep and which we should drop

plot(pca, type = "lines")
```

## Decide which components you would drop and remove them from your data set.
From the plot, the variance of PC4, PC5 and PC6 are all less than 1, which is low enough. But from the variance accounted and the summary of pca, we see that PC6 accounts much lower variance and has much lower proportion of variance, which means PC6 does not contribute a lot. In this case, I would drop only PC6.


## Part II

```{r}
#Now, create a data frame of the transformed data from your pca.

D3 <- data.frame(pca$x)

#Attach the variable "mean_correct" from your original data frame to D3.

D3$mean_correct <- D1$mean_correct

#Now re-run your correlation plots between the transformed data and mean_correct. If you had dropped some components would you have lost important information about mean_correct?

ggpairs(D3, progress = FALSE) 
ggcorr(D3, method = c("everything", "pearson")) 

# From the correlation plots, we known that mean_correct has strong positive correlation with PC1 and PC2. In this case, if we had dropped PC1 and PC2, we would have lost important information. Also, there is a negative correlation between mean_correct and PC6, although the correlation is not too strong, PC6 is the only one holds the negative correlation with mean_correct, we consider not to drop PC6. In this case, if we had dropped PC6, we may also have lost some important information. PC4 has positive correlation with mean_correct but not so strong, in this case, if we had dropped PC3, PC4, PC5, we may not have lost important infomration.
```
## Now print out the loadings for the components you generated:

```{r}
pca$rotation

#Examine the eigenvectors, notice that they are a little difficult to interpret. It is much easier to make sense of them if we make them proportional within each component

loadings <- abs(pca$rotation) #abs() will make all eigenvectors positive

#Now examine your components and try to come up with substantive descriptions of what some might represent?

# PC1 has higher loading score for mean_hint and mean_attempt, which means PC1 might represent how many attempts and efforts did students put in. PC2 has higher loading score for prior_prob_count and prior_percent_correct, which means PC2 might represents the prior abilities and experience for students. PC3 has higher loading score for mean_confidence, which means PC3 might represent students' confidence of their ability.


#You can generate a biplot to help you, though these can be a bit confusing. They plot the transformed data by the first two components. Therefore, the axes represent the direction of maximum variance accounted for. Then mapped onto this point cloud are the original directions of the variables, depicted as red arrows. It is supposed to provide a visualization of which variables "go together". Variables that possibly represent the same underlying construct point in the same direction.  

biplot(pca)

```
# Part III  
Also in this repository is a data set collected from TC students (tc-program-combos.csv) that shows how many students thought that a TC program was related to another TC program. Students were shown three program names at a time and were asked which two of the three were most similar. Use PCA to look for components that represent related programs. Explain why you think there are relationships between these programs.

```{r}
data <- read.csv("tc-program-combos.csv", header = TRUE)
ggcorr(data[,-1], method = c("everything", "pearson"))
data2 <- data[, -1] #remove program name
pca2 <- prcomp(data2, scale. = TRUE)
pca2$sdev
pca2$sdev^2
summary(pca2)
plot(pca2, type = "lines")
loadings2 <- abs(pca2$rotation)
biplot(pca2)

# From biplot, we see that there are many arrows (indicates many programs) almost toward to the same direction (such that Clinical Psychology, Sexuality Women and Gender in Psychology, Counseling Psychology), which means they are highly correlated. In this case, there are relationships between these programs.
```





