---
title: "Assignment 5 - Decision Trees"
author: "Qiyang (Minnie) Lin"
date: "November 9, 2019"
output: html_document
---
For this assignment we will be using data from the Assistments Intelligent Tutoring system. This system gives students hints based on how they perform on math problems. 

#Install & call libraries
```{r}
#install.packages("party") 
#install.packages("rpart")
#install.packages("rpart.plot")

library(rpart)
library(party)
library(rpart.plot)
```

## Part I
```{r}
D1 <- read.csv("intelligent_tutor.csv")
  
```

##Classification Tree
First we will build a classification tree to predict which students ask a teacher for help, which start a new session, or which give up, based on whether or not the student completed a session (D1$complete) and whether or not they asked for hints (D1$hint.y). 
```{r}

c.tree <- rpart(action ~ hint.y + complete, method="class", data=D1) #Notice the standard R notion for a formula X ~ Y

#Look at the error of this tree
printcp(c.tree)
       #CP: complexity parameter. error will go down when there are more splits. relerror: SSE/RMSE, xerror: cross-validation error, xstd. 

#Plot the tree
post(c.tree, file = "tree.ps", title = "Session Completion Action: 1 - Ask teacher, 2 - Start new session, 3 - Give up")
rpart.plot(c.tree)

```
## Part II

#Regression Tree

We want to see if we can build a decision tree to help teachers decide which students to follow up with, based on students' performance in Assistments. We will create three groups ("teacher should intervene", "teacher should monitor student progress" and "no action") based on students' previous use of the system and how many hints they use. To do this we will be building a decision tree using the "party" package. The party package builds decision trees based on a set of statistical stopping rules.

#Visualize our outcome variable "score"
```{r}
hist(D1$score, main = "Score")

```

#Create a categorical outcome variable based on student score to advise the teacher using an "ifelse" statement
```{r}
#"no action" if students score > 0.6, "monitor"if students score between 0.3-0.6, "intervention" if students score < 0.3


#D1$advice <- if(D1$score > 0.7){"no action"
#} else if(D1$score > 0.4 && D1$score < 0.7){"monitor"
#  }else{"intervene"} doesn't work due to length >1 

#D1$advice<- ifelse(D1$score >= 0.7, 
#                   ifelse(D1$score >= 0.4 && D1$score < 0.7,"monitor","no action"),
#                          "intervene") doesn't work with the wrong logic. 


D1$advice <- with(D1, ifelse(score >= 0.3,
                    ifelse(score >= 0.6,"no action","monitor"),
                    "intervene"))

#use with() to avoid metioning df each time. 
                    
```

#Build a decision tree that predicts "advice" based on how many problems students have answered before, the percentage of those problems they got correct and how many hints they required
```{r}
#outcome ~ predictor1+predictor2+predictor3+ect.

score_ctree <- rpart(advice ~ prior_prob_count + prior_percent_correct +hints, method="class", data=D1)
  
```

#Plot tree
```{r}

rpart.plot(score_ctree)
#https://blog.exploratory.io/visualizing-a-decision-tree-using-r-packages-in-explortory-b26d4cb5e71f

printcp(score_ctree)

#plot(score_ctree)

```

Please interpret the tree, which two behaviors do you think the teacher should most closely pay attemtion to?

```{r}
#Interpretation: Depending on whether they had hints over 20, the actions split into two (monitor for 7% and no action for 93%). It means that if a student received hints more or equal to 20 times, 7% chance that teacher shouldn monitor. And if they received hints over 58times, 2% chance that teacher should intervene and 8% should be moniotred. 
#If they received hints less than 20, 93% chance that teacher should ignore. However, when students got less 20 hints, previous correct less then 60% and got more than 1 hints, 19% of them should be monitored.

#how many hints students got and prior percent correct scores will be two behaviors that teachers should most closely pay attention to. (because they are in the higher hierarchy)


```


#Test Tree
Upload the data "intelligent_tutor_new.csv". This is a data set of a differnt sample of students doing the same problems in the same system. We can use the tree we built for the previous data set to try to predict the "advice" we should give the teacher about these new students. 

```{r}
#Upload new data

D2 <- read.csv("intelligent_tutor_new.csv")

#Generate predicted advice using the predict() command for new students based on tree generated from old students

#predict(old, new)
D2$prediction <- predict(score_ctree, D2)

``` 
## Part III
Compare the predicted advice with the actual advice that these students recieved. What is the difference between the observed and predicted results?

```{r}
#advice action in D2 obs.data
D2$advice <- with(D2, ifelse(score >= 0.3,
                    ifelse(score >= 0.6,"no action","monitor"),
                    "intervene"))
#advice from the prediction
D2$advice_Pred<-with(D2, ifelse(prediction >= 0.3,
                    ifelse(prediction >= 0.6,"no action","monitor"),
                    "intervene"))

#compare prediciton action and action.By changing the cutoff score, we want the rel error be higher(it is how we prediction) and xerror (cross validation )be lower
D2$Accuracy<- ifelse(D2$advice==D2$advice_Pred,1, 0)

#calculate the accuracy of the prediction
sum(as.numeric(D2$Accuracy))/200

#prediction advice accuracy is 71.5%


#0.3-0.6 71.5%
#0.7-0.8 47%

```


### To Submit Your Assignment

Please submit your assignment by first "knitting" your RMarkdown document into an html file and then commit, push and pull request both the RMarkdown file and the html file.

