#ASSIGNMENT 5 TEST - 11/26/18
---
title: "Assignment 5 - Decision Trees"
author: "Charles Lang"
date: "November 9, 2016"
output: html_document
---
For this assignment we will be using data from the Assistments Intelligent Tutoring system. This system gives students hints based on how they perform on math problems. 

#Install & call libraries
```{r}

# install.packages("party", "rpart")
#library("party", lib.loc="/Library/Frameworks/R.framework/Versions/3.3/Resources/library")

#FOR SOME REASON, MY PARTY PACKAGE WOULD NOT LOAD, SO I LOADED THESE PACKAGES FIRST
#SUGGESTIONS FROM HERE: https://stackoverflow.com/questions/24687280/error-installing-party-package-in-r


# I RESOLVED MY ISSUE BY POSTING TO STACK OVERFLOW HERE: https://stackoverflow.com/questions/53385377/party-package-installation-failure-and-error-messages?noredirect=1#comment93649496_53385377

install.packages("mvtnorm")
library(mvtnorm)
install.packages("multcomp")
library(multcomp)
install.packages("party")
library("party")
# I did NOT "install from sources the package which needs compilation", which allowed PARTY to finally install!


library(rpart)
library("dplyr", lib.loc="/Library/Frameworks/R.framework/Versions/3.3/Resources/library")
library("ggplot2", lib.loc="/Library/Frameworks/R.framework/Versions/3.3/Resources/library")
library("igraph", lib.loc="/Library/Frameworks/R.framework/Versions/3.3/Resources/library")
library("RColorBrewer", lib.loc="/Library/Frameworks/R.framework/Versions/3.3/Resources/library")
library("tidyr", lib.loc="/Library/Frameworks/R.framework/Versions/3.3/Resources/library")
library("tidyselect", lib.loc="/Library/Frameworks/R.framework/Versions/3.3/Resources/library")
library(factoextra)

install.packages("partykit")
library("partykit")
```

## Part I
```{r}
D1 <- as.data.frame(read.csv("intelligent_tutor.csv"))
```

##Classification Tree
First we will build a classification tree to predict which students ask a teacher for help, which start a new session, or which give up, based on whether or not the student completed a session (D1$complete) and whether or not they asked for hints (D1$hint.y). 
```{r}

c.tree <- rpart(action ~ hint.y + complete, method="class", data=D1) #Notice the standard R notion for a formula X ~ Y

#Look at the error of this tree
printcp(c.tree)

#Plot the tree
post(c.tree, title = "Session Completion Action: 1 - Ask teacher, 2 - Start new session, 3 - Give up", filename = "tree.ps")

# I DON'T KNOW WHAT I AM SUPPOSED TO BE ABLE TO SEE; NOTHING IS APPEARING
```
## Part II

#Regression Tree

We want to see if we can build a decision tree to help teachers decide which students to follow up with, based on students' performance in Assistments. We will create three groups ("teacher should intervene", "teacher should monitor student progress" and "no action") based on students' previous use of the system and how many hints they use. To do this we will be building a decision tree using the "party" package. The party package builds decision trees based on a set of statistical stopping rules.

#Take a look at our outcome variable "score"
```{r}
hist(D1$score)
```

#Create a categorical outcome variable based on student score to advise the teacher using an "ifelse" statement
```{r}
D1$advice <- ifelse(D1$score <=0.4, "intervene", ifelse(D1$score > 0.4 & D1$score <=0.8, "monitor", "no action"))
```

#Build a decision tree that predicts "advice" based on how many problems students have answered before, the percentage of those problems they got correct and how many hints they required
```{r}
score_ctree <- ctree(factor(advice) ~ prior_prob_count + prior_percent_correct + hints, D1)
```

#Plot tree
```{r}
plot(score_ctree)
```

Please interpret the tree, which two behaviors do you think the teacher should most closely pay attemtion to

#### <!-- IN GENERAL, THE TEACHER SHOULD PAY ATTENTION TO STUDENTS WHO NEEDED A LOT OF HINTS (>12), YET STILL HAD A LOW "PERCENT CORRECT" SCORE -->
####  <!-- THE TEACHER SHOULD ALSO WATCH STUDENTS WHO DID NOT USE A LOT OF HINTS, YET HAD A LOW "PRIOR PROBLEM" COUNT -->

  <!-- I DO NOT FEEL COMPLETELY CONFIDENT ABOUT HOW I AM INTERPRETING THE GRAPH BELOW. IS THE DARK BAR INTERVENING? -->
  <!-- I ALSO DO NOT KNOW WHY THE RIGHT-MOST LINE IS NOT CONNECTED TO ANYTHING...IS THIS AN ERROR? IF SOMEONE NEEDED SO MANY HINTS,       
  SHOULDN'T THAT BE A FLAG? -->

  <!-- AM I INTERPRETING THIS CORRECTLY? I AM MOST CONFUSED ABOUT THE BLANK FOR "DO NOTHING" WHEN SOMEONE NEEDS SO MANY HINTS, AND ALSO WHY NODE 8 LOOKS SO DIFFERENT THAN THE REST OF THE BAR GRAPHS. I ALSO DON'T UNDERSTAND THE "P=" IN EACH NODE. IS THIS A PROBABILITY, OR A P-VALUE? AND WHAT DOES THIS MEAN FOR THE DATA-SET SPECIFICALLY? -->

#Test Tree
Upload the data "intelligent_tutor_new.csv". This is a data set of a differnt sample of students doing the same problems in the same system. We can use the tree we built for the previous data set to try to predict the "advice" we should give the teacher about these new students. 

```{r}
#Upload new data

D2 <- as.data.frame(read.csv("intelligent_tutor_new.csv"))

#Generate predicted advice for new students based on tree generated from old students

D2$prediction <- predict(score_ctree, D2)

``` 
## Part III
Compare the predicted advice with the actual advice that these students recieved. What is the difference between the observed and predicted results?

###HOW CAN I COMPARE, WHEN THERE ARE NO ID VARIABLES IN THE 2ND DATASET?
###THE MOST OBVIOUS DIFFERENCE AT FIRST GLANCE, IS THAT THERE IS NO "INTERVENE" FOR D2, MERELY "MONITOR" AND "NO ACTION"
###I'M ALSO NOTICING THAT THE NUMBER OF HINTS IS DIFFERENT FOR ANY GIVEN SCORE IN D2 RELATIVE TO THE HINTS PROVIDED IN D1

#Calculate the sum of the monitor (if noaction==1, ifelse==0, mean)
