---
title: "Principle Component Aanalysis"
output: html_document
author: "Jiasheng (Andy) Yu"
---
## Data
The data you will be using comes from the Assistments online intelligent tutoring system (https://www.assistments.org/). It describes students working through online math problems. Each student has the following data associated with them:

- id
- prior_prob_count: How many problems a student has answered in the system prior to this session
- prior_percent_correct: The percentage of problems a student has answered correctly prior to this session
- problems_attempted: The number of problems the student has attempted in the current session
- mean_correct: The average number of correct answers a student made on their first attempt at problems in the current session
- mean_hint: The average number of hints a student asked for in the current session
- mean_attempt: The average number of attempts a student took to answer a problem in the current session
- mean_confidence: The average confidence each student has in their ability to answer the problems in the current session

## Start by uploading the data
```{r}
D1 <- read.csv("Assistments-confidence.csv")
```

## Create a correlation matrix of the relationships between the variables, including correlation coefficients for each pair of variables/features.

```{r}
#You can install the corrplot package to plot some pretty correlation matrices (sometimes called correlograms)

library(ggplot2)
library(corrplot)
library(GGally)
library(tidyr)
library(dplyr)

library(psych)
ggpairs(D1, 2:8, progress = FALSE) #ggpairs() draws a correlation plot between all the columns you identify by number (second option, you don't need the first column as it is the student ID) and progress = FALSE stops a progress bar appearing as it renders your plot

ggcorr(D1[,-1], method = c("everything", "pearson")) #ggcorr() doesn't have an explicit option to choose variables so we need to use matrix notation to drop the id variable. We then need to choose a "method" which determines how to treat missing values (here we choose to keep everything, and then which kind of correlation calculation to use, here we are using Pearson correlation, the other options are "kendall" or "spearman")

#Study your correlogram images and save them, you will need them later. Take note of what is strongly related to the outcome variable of interest, mean_correct. 
correlogram= cor(D1[,-1])
corrplot(correlogram,method = "color")


```

## Create a new data frame with the mean_correct variable removed, we want to keep that variable intact. The other variables will be included in our PCA.

```{r}
D2 <- D1[,c(-5,-1)]
test_D2<-D1[,c(-1,-5)]

fa.parallel(test_D2,fa='pc',n.iter = 100,show.legend = F,main="Scree plot with parallel analysis")
abline(h = 1,col='red')

```

## Now run the PCA on the new data frame

```{r}
pca <- prcomp(D2,center=T, scale.= TRUE)

pca

```

## Although princomp does not generate the eigenvalues directly for us, we can print a list of the standard deviation of the variance accounted for by each component.

```{r}
pca$sdev

#To convert this into variance accounted for we can square it, these numbers are proportional to the eigenvalue

pca$sdev^2

#A summary of our pca will give us the proportion of variance accounted for by each component

summary(pca)

#We can look at this to get an idea of which components we should keep and which we should drop

plot(pca, type = "lines")
abline(h=1,col='blue')




```

## Decide which components you would drop and remove them from your data set.

**From the above plot graph and parallel graphs, I will remove the fourth,fifth and sixth components, because the first three components are above hline=1(The 1.0 horizontal line is the eigenvalue of the 1 criterion). However, from the summary of pca, we can know the cumulative proportion from PC1 to PC5 can cover above 90% proportion, so I may only remove the sixth component .**

```{r}
temp<-data.frame(pca$x)

```

## Part II

```{r}
#Now, create a data frame of the transformed data from your pca.

D3 <- temp

#Attach the variable "mean_correct" from your original data frame to D3.
D3$mean_correct<-D1$mean_correct

#Now re-run your correlation plots between the transformed data and mean_correct. If you had dropped some components would you have lost important infomation about mean_correct?


ggpairs(D3, progress = FALSE) 

ggcorr(D3, method = c("everything", "pearson")) 

D3_temp<-cor(D3)
corrplot(D3_temp,method='color')

```

**From the different ggpairs graph we can know that mean_correct has strong relationships with PC1 and PC2, so if we keep PC1 and PC2, we will not lose much information. If we use scree test and drop components PC4 and PC5, I think we may not lose much information.**

## Now print out the loadings for the components you generated:

```{r}
pca$rotation

#Examine the eigenvectors, notice that they are a little difficult to interpret. It is much easier to make sense of them if we make them proportional within each component

loadings <- abs(pca$rotation) #abs() will make all eigenvectors positive
loadings
#Now examine your components and try to come up with substantive descriptions of what some might represent?

#You can generate a biplot to help you, though these can be a bit confusing. They plot the transformed data by the first two components. Therefore, the axes represent the direction of maximum variance accounted for. Then mapped onto this point cloud are the original directions of the variables, depicted as red arrows. It is supposed to provide a visualization of which variables "go together". Variables that possibly represent the same underlying construct point in the same direction.  

biplot(pca)

```

**From the loadings, we can know PC1 is more likely to represent how many times and how much effort students put in the exercises. PC2 will represent the degree that students are familiar with the questions.PC3 will represent students' confidence.  **


# Part III  
```{r}
tc_program_df <-read.csv('tc-program-combos.csv')
tc_program_df2<-tc_program_df[,-1]
#pca_test<-fa.parallel(tc_program_df2,fa='pc',n.iter=40)
tc_program_pca <- prcomp(tc_program_df2,scale. = T)
head(summary(pca))
plot(pca,type='lines')
abline(h = 1,col='red')

#tc_program_df2<-cor(tc_program_df2)

head(tc_program_pca$rotation)
tc_loadings<-abs(tc_program_pca$rotation)
biplot(tc_program_pca)



```

**From the graph above, I think it can show clear relationships between different programs. Because from the same direction we can see there are more than one course which mean these two courses are related such as Psychology and Clinical Psychology. **





