---
title: "Principle Component Aanalysis"
output: html_document
---
## Data
The data you will be using comes from the Assistments online intelligent tutoring system (https://www.assistments.org/). It describes students working through online math problems. Each student has the following data associated with them:

- id
- prior_prob_count: How many problems a student has answered in the system prior to this session
- prior_percent_correct: The percentage of problems a student has answered correctly prior to this session
- problems_attempted: The number of problems the student has attempted in the current session
- mean_correct: The average number of correct answers a student made on their first attempt at problems in the current session
- mean_hint: The average number of hints a student asked for in the current session
- mean_attempt: The average number of attempts a student took to answer a problem in the current session
- mean_confidence: The average confidence each student has in their ability to answer the problems in the current session

## Start by uploading the data
```{r}
D1 <- read.csv("Assistments-confidence.csv", header= TRUE)

```

## Create a correlation matrix of the relationships between the variables, including correlation coefficients for each pair of variables/features.
```{r}
library(tidyr)
library(dplyr)
```

```{r}
#You can install the corrplot package to plot some pretty correlation matrices (sometimes called correlograms)

library(ggplot2)
library(GGally)

ggpairs(D1, 2:8, progress = FALSE) #ggpairs() draws a correlation plot between all the columns you identify by number (second option, you don't need the first column as it is the student ID) and progress = FALSE stops a progress bar appearing as it renders your plot

ggcorr(D1[,-1], method = c("everything", "pearson")) #ggcorr() doesn't have an explicit option to choose variables so we need to use matrix notation to drop the id variable. We then need to choose a "method" which determines how to treat missing values (here we choose to keep everything, and then which kind of correlation calculation to use, here we are using Pearson correlation, the other options are "kendall" or "spearman")

#Study your correlogram images and save them, you will need them later. Take note of what is strongly related to the outcome variable of interest, mean_correct. 
```

## Create a new data frame with the mean_correct variable removed, we want to keep that variable intact. The other variables will be included in our PCA.

```{r}
D2 <- select(D1, -mean_correct, -id)

```

## Now run the PCA on the new data frame

```{r}
pca <- prcomp(D2, scale. = TRUE)
```

## Although princomp does not generate the eigenvalues directly for us, we can print a list of the standard deviation of the variance accounted for by each component.

```{r}
pca$sdev

#To convert this into variance accounted for we can square it, these numbers are proportional to the eigenvalue

pca$sdev^2

#A summary of our pca will give us the proportion of variance accounted for by each component

summary(pca)

#We can look at this to get an idea of which components we should keep and which we should drop

plot(pca, type = "lines")
```

## Decide which components you would drop and remove them from your data set.
```{r}
## I would like to drop PC6, because the variance of PC6 is too small compared to the variance of other values, and the previous values are all bigger than it, so PC6 can't fit the whole well.
```


## Part II

```{r}
#Now, create a data frame of the transformed data from your pca.

D3 <- data.frame(pca$x)

#Attach the variable "mean_correct" from your original data frame to D3.

D3$mean_correct <- D1$mean_correct


#Now re-run your correlation plots between the transformed data and mean_correct. If you had dropped some components would you have lost important infomation about mean_correct?

ggpairs(D3, progress = FALSE)
 ggcorr(D3, method = c("everything", "pearson"))

```
## Now print out the loadings for the components you generated:
```{r}
## As shown in the figure, PC1 and PC2 have a very strong correlation, so it is certain that we cannot drop PC1 and PC2. Mean_corre of PC6 is negatively correlated, so if I want to drop PC6, the information lost will be affected. However, it can be seen that the correlation between PC3 and PC5 is minimal and almost negligible, so we will not lose much information.
```


```{r}
pca$rotation

#Examine the eigenvectors, notice that they are a little difficult to interpret. It is much easier to make sense of them if we make them proportional within each component

loadings <- abs(pca$rotation) #abs() will make all eigenvectors positive

#Now examine your components and try to come up with substantive descriptions of what some might represent?

#You can generate a biplot to help you, though these can be a bit confusing. They plot the transformed data by the first two components. Therefore, the axes represent the direction of maximum variance accounted for. Then mapped onto this point cloud are the original directions of the variables, depicted as red arrows. It is supposed to provide a visualization of which variables "go together". Variables that possibly represent the same underlying construct point in the same direction.  

biplot(pca)


```
```{r}
## From the picture above, I think PC1 may represent the effort level of students, because mean_hint and mean attempt have high values. PC2 may represent students' prior knowledge, because prio_precent_correct and prior_prob_count values are higher; PC3 may represent students' confidence, because PC3 has a high mean_confidence score.
```



# Part III  
Also in this repository is a data set collected from TC students (tc-program-combos.csv) that shows how many students thought that a TC program was related to andother TC program. Students were shown three program names at a time and were asked which two of the three were most similar. Use PCA to look for components that represent related programs. Explain why you think there are relationships between these programs.

```{r}
DF1 <- read.csv('tc-program-combos.csv',header = TRUE)
DF2 <- DF1[,-1]
pca2 <- prcomp(DF2, scale. = TRUE)
pca2$sdev
pca2$sdev^2
summary(pca2)
plot(pca2, type = "lines")
pca2$rotation
loadings2 <- abs(pca2$rotation)
biplot(pca2)

```
```{r}
## According to the line diagram of variance, I could see that PC1, PC2, PC3 and PC4 have obvious changes, but after PC4, the variance starts to flatten out, so I may analyze it again by the components after drop PC4. In addition, in the third figure, I found that these programs are correlated to some extent. For example, PC1 seems to get relatively high scores in Education, Arts.Administration and Social, which may be related to programs related to Education. PC4 has high scores in anthropology, linguistics, English and education, so these may all be related to language teaching procedures.
```





