---
title: "Assignment 5 - Decision Trees"
author: "Charles Lang"
date: "November 9, 2016"
output:
  html_document: default
  pdf_document: default
---
For this assignment we will be using data from the Assistments Intelligent Tutoring system. This system gives students hints based on how they perform on math problems. 

#Install & call libraries
```{r}


library(rpart)
library(party)
```

## Part I
```{r}
D1 <- read.csv("~/Desktop/master fall/hudk4050/assignment5/intelligent_tutor.csv")
```

##Classification Tree
First we will build a classification tree to predict which students ask a teacher for help, which start a new session, or which give up, based on whether or not the student completed a session (D1$complete) and whether or not they asked for hints (D1$hint.y). 
```{r}

c.tree <- rpart(action ~ hint.y + complete, method="class", data=D1) #Notice the standard R notion for a formula X ~ Y

#Look at the error of this tree
printcp(c.tree)
```
```{r}
#Plot the tree
post(c.tree, file = "tree.ps", title = "Session Completion Action: 1 - Ask teacher, 2 - Start new session, 3 - Give up")

```
## Part II

#Regression Tree

We want to see if we can build a decision tree to help teachers decide which students to follow up with, based on students' performance in Assistments. We will create three groups ("teacher should intervene", "teacher should monitor student progress" and "no action") based on students' previous use of the system and how many hints they use. To do this we will be building a decision tree using the "party" package. The party package builds decision trees based on a set of statistical stopping rules.

#Visualize our outcome variable "score"
```{r}
hist(D1$score)
```

#Create a categorical outcome variable based on student score to advise the teacher using an "ifelse" statement
```{r}
#set the level of student score:    
#if score between 0-0.4, should be level:"teacher should intervene",in short "intervene"    
#if score between 0.4-0.7, should be level:"teacher should monitor student progress", in short "monitor"        
#if score greater than 0.7, should be level :"no action"   
D1$advice <- ifelse (D1$score >= 0.4, ifelse(D1$score >= 0.7,"no action","monitor"),"intervene" )
```

#Build a decision tree that predicts "advice" based on how many problems students have answered before, the percentage of those problems they got correct and how many hints they required
```{r}
score_ctree <- rpart(factor(advice) ~ prior_prob_count + prior_percent_correct+hints, method="class", data=D1)
```

#Plot tree
```{r}
post(score_ctree, file = "score_tree.ps", title = " three groups:1:teacher should intervene, 2:teacher should monitor student progress,3:no action")
```

Please interpret the tree, which two behaviors do you think the teacher should most closely pay attemtion to?
i think the following two types of students that the teacher should most closely pay attemtion :       
students who require hints greater than 12.5       
students who require hints smaller than 12.5 and greater than 0.5, percentage of those problems they got correct smaller than 40%.      
for these two types of students, they are mostly the student need to be monitored or intervened.   
#Test Tree    
Upload the data "intelligent_tutor_new.csv". This is a data set of a differnt sample of students doing the same problems in the same system. We can use the tree we built for the previous data set to try to predict the "advice" we should give the teacher about these new students. 

```{r}
#Upload new data

D2 <- read.csv("~/Desktop/master fall/hudk4050/assignment5/intelligent_tutor_new.csv")

#Generate predicted advice using the predict() command for new students based on tree generated from old students

D2$prediction <- predict(score_ctree,D2,type = "class")
``` 
## Part III
Compare the predicted advice with the actual advice that these students recieved. What is the difference between the observed and predicted results?
```{r}
library(dplyr)
D2$advice <- ifelse (D2$score >= 0.4, ifelse(D2$score >= 0.7,'no action','monitor'),'intervene' )
comb <- select(D2,advice,prediction)
confMat <- table(comb)
confMat
accuracy <- sum(diag(confMat))/sum(confMat)
accuracy
```
the accuracy is 42.5% which means the prediction only make the 42.5% corret predicts of total students. from the new data set we can see that the score of all student is 1, so all student under the level of no action, but by prediction, there are 42.5% of student should under level of no action.    
for prediction value, the student that should be monitored is 14%, but the actual value is 0%.
for prediction value, the student that should be intervened is 42.5%, but the actual value is 0%.
the difference between prediction and actual value is huge.

### To Submit Your Assignment

Please submit your assignment by first "knitting" your RMarkdown document into an html file and then commit, push and pull request both the RMarkdown file and the html file.

