---
title: "Assignment 5 - Decision Trees"
author: "Charles Lang"
date: "November 9, 2016"
output: html_document
---
For this assignment we will be using data from the Assistments Intelligent Tutoring system. This system gives students hints based on how they perform on math problems. 

#Install & call libraries
```{r}
install.packages("party", "rpart")

library(rpart)
library(party)
```

#Upload Data
```{r}
D1 <- read.table("intelligent_tutor.csv", sep = ",", header = TRUE)
```

##Classification Tree
First we will build a classification tree to predict which students ask a teacher for help, which start a new session, or which give up, based on whether or not the student completed a session (D1$complete) and whether or not they asked for hints (D1$hint.y). 
```{r}

c.tree <- rpart(action ~ hint.y + complete, method="class", data=D1) #Notice the standard R notion for a formula X ~ Y

#Look at the error of this tree
printcp(c.tree)

#Plot the tree
post(c.tree, file = "tree.ps", title = "Session Completion Action: 1 - Ask teacher, 2 - Start new session, 3 - Give up")

```
#Regression Tree

We want to see if we can build a decision tree to help teachers decide which students to follow up with, based on students' performance in Assistments. We will create three groups ("teacher should intervene", "teacher should monitor student progress" and "no action") based on students' previous use of the system and how many hints they use. To do this we will be building a decision tree using the "party" package. The party package builds decision trees based on a set of statistical stopping rules.

#Take a look at our outcome variable "score"
```{r}
hist(D1$score)
```

#Create a categorical outcome variable based on student score to advise the teacher using an "ifelse" statement
```{r}
D1$advice <- ifelse(D1$score <=0.4, "intervene", ifelse(D1$score > 0.4 & D1$score <=0.8, "monitor", "no action"))
```

#Build a decision tree that predicts "advice" based on how many problems students have answered before, the percentage of those problems they got correct and how many hints they required
#what is the different between the ctree command and rpart?
```{r}
score_ctree <- ctree(factor(advice) ~ prior_prob_count + prior_percent_correct + hints, D1)
```

#Plot tree
```{r}
plot(score_ctree)
```

Please interpret the tree, which two behaviors do you think the teacher should most closely pay attemtion to?
A: Since we can see that for those studnets in the branch of "ask for hint" usually needs higher monitoring and those who got a "percent correct" lower than 12 need more intervention. Thus, teacher should focus on those who asked for hint and those who have a lower prior percent correct than 12.


#Test Tree
Upload the data "intelligent_tutor_new.csv" and use the predict function (D2$prediction <- predict(score_ctree, D2)) to predict the assignments of the new data set. What is the error rate on your predictions of the new data? 

upload the csv to d2
```{r}
D2 <- read.table("intelligent_tutor_new.csv", sep = ",", header = TRUE)

```

use the given code to predict:
```{r}
D2$prediction <- predict(score_ctree, D2)
```

due to our early suggestion, we should intervene if score <=0.4, moniter if 0.4<=score<=0.8, and take no action if score>=0.8
```{r}
D2$advice <- ifelse(D2$score <=0.4, "intervene", ifelse(D2$score > 0.4 & D2$score <=0.8, "monitor", "no action"))
#now find out under how many cases that advice dismatch with the preditiction
D2$dismatch <- ifelse(D2$prediction != D2$advice, 1, 0)
#count the number of dismatch
n=sum(D2$dismatch)
#compute the error rate
errorrate=n/200

```

