---
title: "Assignment 5 - Decision Trees"
author: "Charles Lang"
date: "November 9, 2016"
output: html_document
---
For this assignment we will be using data from the Assistments Intelligent Tutoring system. This system gives students hints based on how they perform on math problems. 

#Install & call libraries
```{r}
##install.packages("party", "rpart")

library(rpart)
library(party)
```

#Upload Data
```{r}
D1 <- read.table("intelligent_tutor.csv", sep = ",", header = TRUE)
```

##Classification Tree
First we will build a classification tree to predict which students ask a teacher for help, which start a new session, or which give up, based on whether or not the student completed a session (D1$complete) and whether or not they asked for hints (D1$hint.y). 
```{r}

c.tree <- rpart(action ~ hint.y + complete, method="class", data=D1) #Notice the standard R notion for a formula X ~ Y

#Look at the error of this tree
printcp(c.tree)

#Plot the tree
post(c.tree, file = "tree.ps", title = "Session Completion Action: 1 - Ask teacher, 2 - Start new session, 3 - Give up")

```
#Regression Tree

We want to see if we can build a decision tree to help teachers decide which students to follow up with, based on students' performance in Assistments. We will create three groups ("teacher should intervene", "teacher should monitor student progress" and "no action") based on students' previous use of the system and how many hints they use. To do this we will be building a decision tree using the "party" package. The party package builds decision trees based on a set of statistical stopping rules.

#Take a look at our outcome variable "score"
```{r}
hist(D1$score)
```

#Create a categorical outcome variable based on student score to advise the teacher using an "ifelse" statement
```{r}
D1$advice <- ifelse(D1$score <=0.4, "intervene", ifelse(D1$score > 0.4 & D1$score <=0.8, "monitor", "no action")
```
D2$correct <- ifelse(D2$prediction == "no action", 1, 0)


#Build a decision tree that predicts "advice" based on how many problems students have answered before, the percentage of those problems they got correct and how many hints they required
```{r}
score_ctree <- ctree(factor(advice) ~ prior_prob_count + prior_percent_correct + hints, D1)
```

#Plot tree
```{r}
plot(score_ctree)
```

Please interpret the tree, which two behaviors do you think the teacher should most closely pay attemtion to?

The tree tells us that starting from how many number of hints student requested in the current session, if students ask for a lot of hints ( more than 12) then teachers should not have no action, and should definitely intervene/monitor 

If the number of hints is smaller than 12 but bigger than 0, then depending on the prior percentage of correct answer students get, teachers will decide whether to intervene or not 

If the students did not ask for any hints, then depending on the number of problems students has done prior to the session, teachers will decide on the intervention. If students did more than 85 problems then teachers without any hints then teachers should not intervene/monitor 

The tree tells us that a teacher should most closely pay attention to the number of hints the student requested in the current session, and the prior percentage of correct problems students got. These 2 variables will tell a teacher the most information about intervention 



#Test Tree
Upload the data "intelligent_tutor_new.csv" and use the predict function (D2$prediction <- predict(score_ctree, D2)) to predict the assignments of the new data set. What is the error rate on your predictions of the new data? 


D2 <- read.table("intelligent_tutor_new.csv", sep = ",", header = TRUE)

D2$prediction <- predict(score_ctree, D2)
### Create the new correct variable to compare D2 prediction with D2 score 

D2$correct <- ifelse(D2$prediction == "no action", 1, 0)

### Calculate the mean of D2$correct

summary(D2$correct)

### The percentage of correct is 58%. That means the error rate is 0.42




