---
title: "Assignment 5 - Decision Trees"
author: "Charles Lang"
date: "November 9, 2016"
output: html_document
---
For this assignment we will be using data from the Assistments Intelligent Tutoring system. This system gives students hints based on how they perform on math problems. 

#Install & call libraries
```{r}
install.packages("party", "rpart")

library(rpart)
library(party)
```

#Upload Data
```{r}
D1 <- read.table("intelligent_tutor.csv", sep = ",", header = TRUE)
```

##Classification Tree
First we will build a classification tree to predict which students ask a teacher for help, which start a new session, or which give up, based on whether or not the student completed a session (D1$complete) and whether or not they asked for hints (D1$hint.y). 
```{r}

c.tree <- rpart(action ~ hint.y + complete, method="class", data=D1) 

#Notice the standard R notion for a formula X ~ Y

#Look at the error of this tree
printcp(c.tree)

#Plot the tree
post(c.tree, file = "tree.ps", title = "Session Completion Action: 1 - Ask teacher, 2 - Start new session, 3 - Give up")

```
#Regression Tree

We want to see if we can build a decision tree to help teachers decide which students to follow up with, based on students' performance in Assistments. We will create three groups ("teacher should intervene", "teacher should monitor student progress" and "no action") based on students' previous use of the system and how many hints they use. To do this we will be building a decision tree using the "party" package. The party package builds decision trees based on a set of statistical stopping rules.

#Take a look at our outcome variable "score"
```{r}
hist(D1$score)
```

#Create a categorical outcome variable based on student score to advise the teacher using an "ifelse" statement
```{r}
D1$advice <- ifelse(D1$score <=0.4, "intervene", ifelse(D1$score > 0.4 & D1$score <=0.8, "monitor", "no action"))
```

#Build a decision tree that predicts "advice" based on how many problems students have answered before, the percentage of those problems they got correct and how many hints they required
```{r}
score_ctree <- ctree(factor(advice) ~ prior_prob_count + prior_percent_correct + hints, D1)
```

#Plot tree
```{r}
plot(score_ctree)
```

Please interpret the tree, which two behaviors do you think the teacher should most closely pay attemtion to?

based on score_ctree.pdf plot, students who ask for more than 12 hints and students who ask for less than 12 hints but also have prior percent correct less than .629

#Test Tree
Upload the data "intelligent_tutor_new.csv" and use the predict function (D2$prediction <- predict(score_ctree, D2)) to predict the assignments of the new data set. What is the error rate on your predictions of the new data? 

```{r}
D2 <- read.table("intelligent_tutor_new.csv", sep = ",", header = TRUE)

D2$prediction <- predict(score_ctree, D2)

D2$advice <- ifelse(D2$score <=0.4, "intervene", ifelse(D2$score > 0.4 & D2$score <=0.8, "monitor", "no action"))

#figure out how to do ifelse statement - if this then 1, otherwise 0, add up everything divide by total number 

#what variables should be here? what does the error rate measure against? 
#wrong one 
c.treenew <- rpart(action ~ score + hints, method="class", data=D2) 

#Error in eval(expr, envir, enclos) : object 'action' not found

score_ctree <- ctree(factor(advice) ~ prior_prob_count + prior_percent_correct + hints, D2)

#Look at the error of this tree
printcp(c.tree)

Classification tree:
rpart(formula = action ~ hint.y + complete, data = D1, method = "class")

Variables actually used in tree construction:
[1] complete hint.y  

Root node error: 250/378 = 0.66138

n= 378 

     CP nsplit rel error xerror     xstd
1 0.052      0     1.000  1.128 0.033851
2 0.012      1     0.948  1.036 0.036119
3 0.010      2     0.936  1.016 0.036513

table(D2$prediction, D2$advice)

# no action
 # no action       200

#scores are all 1, so predctions and actions are all none. Did I do something wrong, or was this the right end result? 
  
#Plot the tree
post(c.treenew, file = "treenew.ps", title = "Session Completion Action: 1 - Ask teacher, 2 - Start new session, 3 - Give up")
```