---
title: "Assignment 5 - Decision Trees"
author: "Meijuan Zeng"
output: html_document
---
For this assignment we will be using data from the Assistments Intelligent Tutoring system. This system gives students hints based on how they perform on math problems. 

id - student id prior_prob_count - The number of problems a student has done in the system prior to the surrent session
score - The score the student achieved in the current session
hints - The number of hints the student requested in the current session
hint.y - Whether or not the student asked for hints in the current session
complete - Whether or not the student completed the cirrent session
action - The action suggested by the system to a teacher about a given student based on their performance
#Install & call libraries
```{r}
install.packages("party", "rpart")

library(rpart)
library(party)
```

## Part I
```{r}
D1 <- read.csv("intelligent_tutor.csv", header = TRUE)
```

##Classification Tree
First we will build a classification tree to predict which students ask a teacher for help, which start a new session, or which give up, based on whether or not the student completed a session (D1$complete) and whether or not they asked for hints (D1$hint.y). 
```{r}

c.tree <- rpart(action ~ hint.y + complete, method="class", data=D1) #Notice the standard R notion for a formula X ~ Y

#Look at the error of this tree
printcp(c.tree)

#Plot the tree
post(c.tree, file = "tree.ps", title = "Session Completion Action: 1 - Ask teacher, 2 - Start new session, 3 - Give up")

```
## Part II

#Regression Tree

We want to see if we can build a decision tree to help teachers decide which students to follow up with, based on students' performance in Assistments. We will create three groups ("teacher should intervene", "teacher should monitor student progress" and "no action") based on students' previous use of the system and how many hints they use. To do this we will be building a decision tree using the "party" package. The party package builds decision trees based on a set of statistical stopping rules.

#Take a look at our outcome variable "score"
```{r}
hist(D1$score)
```

#Create a categorical outcome variable based on student score to advise the teacher using an "ifelse" statement
```{r}
D1$advice <- ifelse(D1$score <=0.4, "intervene", ifelse(D1$score > 0.4 & D1$score <=0.8, "monitor", "no action"))
```

#Build a decision tree that predicts "advice" based on how many problems students have answered before, the percentage of those problems they got correct and how many hints they required
```{r}
score_ctree <- ctree(factor(advice) ~ prior_prob_count + prior_percent_correct + hints, D1)
```

#Plot tree
```{r}
plot(score_ctree)
```

Please interpret the tree, which two behaviors do you think the teacher should most closely pay attemtion to?

#The types of advice given by teachers depends on how many problems students have answered before, the percentage of those problems they got correct and how many hints they required. For students who require no hints, there are 145 students who answered less than 85 questions and 76 who answered more than 85 questions. For students who require hints, there are 46 who require more than 12 hints; 45 who require less than 12 hints and score higher than 0.629 and 66 who require less than 12 hints but score lower than 0.629. Teacher should pay most closely attention to students who require more than 12 hints, and less than 12 hints but score lower than 0.629. 

#Test Tree
Upload the data "intelligent_tutor_new.csv". This is a data set of a differnt sample of students doing the same problems in the same system. We can use the tree we built for the previous data set to try to predict the "advice" we should give the teacher about these new students. 

```{r}
#Upload new data

D2 <- read.csv("intelligent_tutor_new.csv", header = TRUE)

#Generate predicted advice for new students based on tree generated from old studentsq
D2$prediction <- predict(score_ctree, D2)
``` 
## Part III
Compare the predicted advice with the actual advice that these students recieved. What is the difference between the observed and predicted results?

```{r}
D2$actual <- ifelse(D2$prediction == "no action", 1, 0)
sum(D2$actual)
#difference: 200 - 116 = 84
#accuracy: 116 / 200 = 0.58
```

