---
title: "Principle Component Aanalysis"
output: html_document
author: Xingyi Xie
---
## Data
The data you will be using comes from the Assistments online intelligent tutoring system (https://www.assistments.org/). It describes students working through online math problems. Each student has the following data associated with them:

- id
- prior_prob_count: How many problems a student has answered in the system prior to this session
- prior_percent_correct: The percentage of problems a student has answered correctly prior to this session
- problems_attempted: The number of problems the student has attempted in the current session
- mean_correct: The average number of correct answers a student made on their first attempt at problems in the current session
- mean_hint: The average number of hints a student asked for in the current session
- mean_attempt: The average number of attempts a student took to answer a problem in the current session
- mean_confidence: The average confidence each student has in their ability to answer the problems in the current session

## Start by uploading the data
```{r}
D1 <- read.csv("Assistments-confidence.csv")

```

## Create a correlation matrix of the relationships between the variables, including correlation coefficients for each pair of variables/features.

```{r}
#You can install the corrplot package to plot some pretty correlation matrices (sometimes called correlograms)

library(ggplot2)
library(GGally)

ggpairs(D1, 2:8, progress = FALSE) #ggpairs() draws a correlation plot between all the columns you identify by number (second option, you don't need the first column as it is the student ID) and progress = FALSE stops a progress bar appearing as it renders your plot

ggcorr(D1[,-1], method = c("everything", "pearson")) #ggcorr() doesn't have an explicit option to choose variables so we need to use matrix notation to drop the id variable. We then need to choose a "method" which determines how to treat missing values (here we choose to keep everything, and then which kind of correlation calculation to use, here we are using Pearson correlation, the other options are "kendall" or "spearman")

#Study your correlogram images and save them, you will need them later. Take note of what is strongly related to the outcome variable of interest, mean_correct. 
```
Note:mean_correct correlates with mean_hint.

## Create a new data frame with the mean_correct variable removed, we want to keep that variable intact. The other variables will be included in our PCA.

```{r}
D2 <- D1[,-5]

```

## Now run the PCA on the new data frame

```{r}
pca <- prcomp(D2, scale. = TRUE)
```

## Although princomp does not generate the eigenvalues directly for us, we can print a list of the standard deviation of the variance accounted for by each component.

```{r}
pca$sdev

#To convert this into variance accounted for we can square it, these numbers are proportional to the eigenvalue

pca$sdev^2

#A summary of our pca will give us the proportion of variance accounted for by each component

summary(pca)

#We can look at this to get an idea of which components we should keep and which we should drop

plot(pca, type = "lines")
```

## Decide which components you would drop and remove them from your data set.
Note: remove PC7

## Part II

```{r}
#Now, create a data frame of the transformed data from your pca.

D3 <- predict(pca)

#Attach the variable "mean_correct" from your original data frame to D3.
D4 <- as.data.frame(cbind(D3,D1$mean_correct))

#Now re-run your correlation plots between the transformed data and mean_correct. If you had dropped some components would you have lost important infomation about mean_correct?
ggpairs(D4, 1:8, progress = FALSE)
ggcorr(D4, method = c("everything", "pearson")) 


```

## Answer: PCA6 negtively correlates with mean_correct and PCA3 positively correlates with mean_correct.

##Now print out the loadings for the components you generated:

```{r}
pca$rotation

#Examine the eigenvectors, notice that they are a little difficult to interpret. It is much easier to make sense of them if we make them proportional within each component

loadings <- abs(pca$rotation) #abs() will make all eigenvectors positive
```

#Now examine your components and try to come up with substantive descriptions of what some might represent?

## Answer: From this we found mean_hint contributes 78.06% variances of PC6, and prior_percent_correct contributes 78.12% variances of PC3. Therefore, we draw a conclusion that pc6 represented mean_hint and it is the most important factor to predict mean_correct.

```{r}
biplot(pca)

library(ggbiplot)
library(factoextra)
library(FactoMineR)

get_pca_var(pca)$contrib#contrib=(var.cos2 * 100) / (total cos2 of the PC component)
fviz_eig(pca,addlabels = TRUE)+labs(title = "Principal component analysis")
pca1 <- fviz_pca_biplot(pca,  palette = "jco", 
                addEllipses = TRUE, label = "var",
                col.var = "black", repel = TRUE)
pca1

```

# Part III  
Also in this repository is a data set collected from TC students (tc-program-combos.csv) that shows how many students thought that a TC program was related to another TC program. Students were shown three program names at a time and were asked which two of the three were most similar. Use PCA to look for components that represent related programs. Explain why you think there are relationships between these programs.

```{r}
D1 <- read.csv("tc-program-combos.csv")
```

```{r}
pca <- prcomp(D1[,-1],rank=10)
pca <- prcomp(D1[,-1],rank=2)
pca <- prcomp(D1[,-1],rank=1)
pca <- prcomp(D1[,-1],rank=5)

```


```{r}
pca$sdev

pca$sdev^2

summary(pca)

plot(pca, type = "lines")
```


```{r}
D3 <- as.data.frame(predict(pca))

ggcorr(D1[,-1], method = c("everything", "pearson")) 


```


```{r message=FALSE, warning=FALSE}

loadings <- as.data.frame(abs(pca$rotation)) #abs() will make all eigenvectors positive
ggcorr(loadings, method = c("everything", "pearson")) #pc5~pc2
ggpairs(loadings, 1:5, progress = FALSE)#PC1~PC3,PC2~PC3,PC2~PC5

library("corrplot")
corrplot(get_pca_var(pca)$cos2, is.corr=FALSE)#visualization,contribution, BUT NO FINDING 

which(loadings$PC1>0.25)
rownames(loadings[which(loadings$PC1>0.25),])
rownames(loadings[which(loadings$PC1>0.2),])

rownames(loadings[which(loadings$PC2>0.25),])
rownames(loadings[which(loadings$PC2>0.2),])

rownames(loadings[which(loadings$PC3>0.25),])
rownames(loadings[which(loadings$PC3>0.2),])

rownames(loadings[which(loadings$PC5>0.25),])
rownames(loadings[which(loadings$PC5>0.2),])

```


## Conclusion: 

- Firstly, we found PC1 is correlated with PC3, and from the contribution table "loading", we found "Change.Leadership" and "Economics.and.Education" are clustered into PC1 and both of them contribute more than 25 percent variances to component 1. Besides, "Arts.Administration", "History", "Politics" and "School.Principals" also contribute more than 20 percent variances to PC1. In my opinion, these programs share something with leadership and regulations, and they seem to be more serious.  
- For PC3, we found  "Clinical.Psychology", "Neuroscience","Psychology" all contribute more than 25 percent variances to PC3, and each of"Physiology","Behavior.Analysis",Cognitive.Science", "Creative.Technologies" also contribute more tahn 20 percent to PC3. From my point of view, these programs share similarity of cognitive behaviour science. 
- For PC5, these programs are more likely to be math related, which requires statistics and compution. 
- However, we found correlation between PC2 and PC3, and we also found correlation between PC2 and PC3. Therefore, considering we didn't find the similarity among programs of "Linguistics", "Creative.Technologies", "Design.and.Development.of.Digital.Games", I think these programs were a mixed version of PC3 and PC5, which means PC2 represented both programs of cognitive behavior science and mathmatics.


```{r message=FALSE, warning=FALSE}

biplot(pca)
library(ggbiplot)
library(factoextra)
library(FactoMineR)


fviz_eig(pca,addlabels = TRUE)+labs(title = "Principal component analysis")
pca2 <- fviz_pca_biplot(pca)
pca2
```



