---
title: "Assignment 5 - Decision Trees"
author: "Charles Lang"
date: "November 9, 2016"
output: html_document
---
For this assignment we will be using data from the Assistments Intelligent Tutoring system. This system gives students hints based on how they perform on math problems. 

#Install & call libraries
```{r}
#install.packages("party", "rpart","rattle", "RColorBrewer")
#install.packages("party", "rpart")
library(rpart)
library(party)
library(rattle)
library(RColorBrewer)

```

## Part I
```{r}
D1 <- read.csv("~/R/HUDK/assignment5/intelligent_tutor.csv")
```

##Classification Tree
First we will build a classification tree to predict which students ask a teacher for help, which start a new session, or which give up, based on whether or not the student completed a session (D1$complete) and whether or not they asked for hints (D1$hint.y). 
```{r}

c.tree <- rpart(action ~ hint.y + complete, method="class", data=D1) #Notice the standard R notion for a formula X ~ Y

#Look at the error of this tree
printcp(c.tree)

#Plot the tree
post(c.tree, file = "tree.ps", title = "Session Completion Action: 1 - Ask teacher, 2 - Start new session, 3 - Give up")

fancyRpartPlot(c.tree, caption = NULL)

```
## Part II

#Regression Tree

We want to see if we can build a decision tree to help teachers decide which students to follow up with, based on students' performance in Assistments. We will create three groups ("teacher should intervene", "teacher should monitor student progress" and "no action") based on students' previous use of the system and how many hints they use. To do this we will be building a decision tree using the "party" package. The party package builds decision trees based on a set of statistical stopping rules.

#Visualize our outcome variable "score"
```{r}
hist(D1$score)
```

#Create a categorical outcome variable based on student score to advise the teacher using an "ifelse" statement

It seems that the distribution of score is "tri-modal". That is, there are 3 groups of students gathered around 3 different score values. We will reccoment "advice" for those students in the middle range, with seems to be bewwwn .3 and .6.

```{r}
D1$advice <- as.factor(ifelse(.3 <D1$score & D1$score <= .6, 1, 0))
```

#Build a decision tree that predicts "advice" based on how many problems students have answered before, the percentage of those problems they got correct and how many hints they required
```{r}
score_ctree <- rpart(advice ~ prior_prob_count + prior_percent_correct + hints, method="class", data=D1)
  
```

#Plot tree
```{r}
fancyRpartPlot(score_ctree, caption = NULL)
```

Please interpret the tree, which two behaviors do you think the teacher should most closely pay attemtion to?

#Test Tree
Upload the data "intelligent_tutor_new.csv". This is a data set of a differnt sample of students doing the same problems in the same system. We can use the tree we built for the previous data set to try to predict the "advice" we should give the teacher about these new students. 

ANSWER:

1) Students that ask for any hints. The first behavior teachers should pay attention to is students that ask for any hints. This category includes 42% of the population, and among these, 48% should receive advice.

2) Students whose percentage of prior correct answers is below .6 (among students that asked for hints).Among those students that asked for hints, paying attention to their prior correct answer percentage provides adittional information for teachers about which students are likely to benefit from receiving advice. Among students with prior correct answer rate below .6 AND who have asked for at least one hint, 60% should receive advice. This "dual" category ("at leas 1 hint" + "correct answer rate below .6") includes 37.7% of the total students who should receive advice ( = .22*.6/.35)  

```{r}
#Upload new data

D2 <- read.csv("~/R/HUDK/assignment5/intelligent_tutor_new.csv")
D2$advice <- ifelse(.3 <D2$score & D2$score <= .6, 1, 0)

#Generate predicted advice using the predict() command for new students based on tree generated from old students

D2$prediction <- predict(score_ctree, D2, type = "class")
  
``` 

## Part III

Compare the predicted advice with the actual advice that these students recieved. What is the difference between the observed and predicted results?

```{r}
#Upload new data

D2 <- read.csv("~/R/HUDK/assignment5/intelligent_tutor_new.csv")
D2$advice <- ifelse(D2$score == 1, 0, 1) # Anna told me that for this variable, students coded as "1" should not receive advice, while students coded "0" should (I find this actually very weired since everyone in the dataset has score =1)

#Generate predicted advice using the predict() command for new students based on tree generated from old students

D2$prediction <- predict(score_ctree, D2, type = "class")

# let's build a two-way percentages table to see what we predicted right and wrong
mytable <- table(as.factor(D2$advice),as.factor(D2$prediction))
prop.table(mytable)
``` 

It seems that the new data is very different from our training data, in the sense that in the training data we had 35% of students needing advice, while in the new data 0% of students needed advice. On the other hand, student inputs seem to be more stable (prior_percent_correct, hints, prior_prob_count), leading to a percentage of predicted "in-need-of-adcive-students" of 27%. Overall, our recommendation was "correct" for the 73% of students for whom we did not recommend advice. This is not that much of a good model though, when we compare to simply recommending the "most usual" state (no advice) - if we had done that, we would have got 100% correct recommendations.    






### To Submit Your Assignment

Please submit your assignment by first "knitting" your RMarkdown document into an html file and then commit, push and pull request both the RMarkdown file and the html file.

