---
title: "Principle Component Analysis"
author: "Nicole Schlosberg"
date: "11/19/2020"
output: html_document
---

Data:  The data comes from the Assistments online intelligent tutoring system (https://www.assistments.org/). It describes students working through online math problems. Each student has the following data associated with them:


## Part I
Uploading the data
```{r}
D1 <- read.csv("Assistments-confidence.csv", header=TRUE)
```

Create a correlation matrix of the relationships between the variables, including correlation coefficients for each pair of variables/features.

```{r}
library(ggplot2)
library(GGally)
library(corrplot)

ggpairs(D1, 2:8, progress = FALSE) #ggpairs() draws a correlation plot between all the columns you identify by number (second option, you don't need the first column as it is the student ID) and progress = FALSE stops a progress bar appearing as it renders your plot

ggcorr(D1[,-1], method = c("everything", "pearson")) #ggcorr() doesn't have an explicit option to choose variables so we need to use matrix notation to drop the id variable. We then need to choose a "method" which determines how to treat missing values (here we choose to keep everything, and then which kind of correlation calculation to use, here we are using Pearson correlation, the other options are "kendall" or "spearman")

#Note of what is strongly related to the outcome variable of interest, mean_correct. 
```

Create a new data frame with the mean_correct variable removed, we want to keep that variable intact. The other variables will be included in our PCA.

```{r}
library(dplyr)

D2 <- select(D1,-id,-mean_correct)
```

Now run the PCA on the new data frame

```{r}
pca <- prcomp(D2, scale. = TRUE)
```

Although princomp does not generate the eigenvalues directly for us, we can print a list of the standard deviation of the variance accounted for by each component.

```{r}
pca$sdev

#To convert this into variance accounted for we can square it, these numbers are proportional to the eigenvalue
pca$sdev^2

#A summary of our pca will give us the proportion of variance accounted for by each component
summary(pca)

#We can look at this to get an idea of which components we should keep and which we should drop
plot(pca, type = "lines")
```

Decide which components you would drop and remove them from your data set.

ANSWER: PC5 and PC6 would be components to drop. PC5 only has 12.20% of variance, meaning it only represents 12.20% of the variance. PC6 only has 8.93% of the variance, which means it only represents 8.93% of the variance. Since it is such small amount of the variance it can be removed.


## Part II

```{r}
#Now create a data frame of the transformed data from your pca.
D3 <- data.frame(pca$x)

#Attach the variable "mean_correct" from your original data frame to D3.
D3 <- data.frame(D3,D1$mean_correct)
 
#Now re-run your correlation plots between the transformed data and mean_correct. If you had dropped some components would you have lost important information about mean_correct?
ggcorr(D3, method = c("everything", "pearson")) 
```

If you had dropped some components would you have lost important information about mean_correct?

ANSWER: You would lose important information about mean_correct. Components with some small variance representation still contribute some information even if small. PC6 which only makes up 8.93% of the variance is strongly correlated to mean_correct. This means that even though it is the best option to drop, we risk dropping something that correlates with the thing we are interested in.

Now print out the loadings for the components you generated:

```{r}
library(ggbiplot)
library(tidyr)

pca$rotation

#Examine the eigenvectors, notice that they are a little difficult to interpret. It is much easier to make sense of them if we make them proportional within each component
loadings <- abs(pca$rotation) #abs() will make all eigenvectors positive

#Now examine your components and try to come up with substantive descriptions of what some might represent?
L1 <- as_tibble(loadings)
labels <- c("prior_prob_count","prior_percent_correct","problems_attempted","mean_hint","mean_attempt","mean_confidence") 
L2 <- cbind(labels,L1)
L3 <- L2 %>% mutate(PC1 = PC1/sum(PC1)) %>% mutate(PC2 = PC2/sum(PC2)) %>% mutate(PC3 = PC3/sum(PC3)) %>% mutate(PC4 = PC4/sum(PC4)) %>% mutate(PC5 = PC5/sum(PC5)) %>%  mutate(PC6 = PC6/sum(PC6)) %>% print

#You can generate a biplot to help you, though these can be a bit confusing. They plot the transformed data by the first two components. Therefore, the axes represent the direction of maximum variance accounted for. Then mapped onto this point cloud are the original directions of the variables, depicted as red arrows. It is supposed to provide a visualization of which variables "go together". Variables that possibly represent the same underlying construct point in the same direction.  
ggbiplot(pca)
#ggbiplot(pca,choices=c(3,4))
#ggbiplot(pca,choices=c(5,6))
```

Now examine your components and try to come up with substantive descriptions of what some might represent?

ANSWER: Having the most variance of a specific category means that the component contributes the most to its variance. PC1 has the most variance in mean_hint (30.22% of variance), mean_attempt (25.86% of variance), and problems_attempted (21.74% of the variance). PC2 contributes the most to prior_percent_correct (44.68% of variance), prior_prob_count (25.08% of variance), and problems_attempted (17.34% of variance). PC3 contributes the most to mean_confidence (45.79% of variance), prior_prob_count (22.10% of variance), and problems_attempted (20.06% of variance). PC4 contributes the most to prior_prob_count (31.52% of variance) and mean_confidence (22.61% of variance). PC5 contributes the most to problems_attempted (30.39% of variance) and mean_attempt (35.77% of variance). PC6 only contributes the most to mean_hint (35.61% of variance). 


## Part III  

Also in this repository is a data set collected from TC students (tc-program-combos.csv) that shows how many students thought that a TC program was related to another TC program. Students were shown three program names at a time and were asked which two of the three were most similar. 

```{r}
library(ggplot2)
library(GGally)
library(dplyr)

R1 <- read.csv("tc-program-combos.csv",header=TRUE)

#Organize the data
R3 <- R1
rownames(R3)<-R3$program
R3 <- select(R3, -program)
R3 <- R3[order(rownames(R3)),]
R3 <- R3[,sort(colnames(R3))]

#PCA on data
pca3 <- prcomp(R3, scale. = TRUE)
pca3$sdev
pca3$sdev^2
summary(pca3)

plot(pca3, type = "lines")
 
#pca3$rotation
loadings3 <- abs(pca3$rotation) 
```

Use PCA to look for components that represent related programs. Explain why you think there are relationships between these programs (explain the meaning of the components).

ANSWER: Based off of the percentages of variance of the PCs, I have concluded that yes many of the PCs correspond to related classes. Below are the make up of the first 6 components, which compose 38.05% (cumulative proportion) of all the variance just within those 6 components. The remaining 61.95% is spread across the other 59 components, thus with 38.05% within 6 components and 61.95% spread across 59 components, the first components hold larger proportions (decreasing as they move away from PC1 towards PC67) of the variance. 

PC1: Change.Leadership (27.70%), Economics.and.Education (23.09%), Education.Policy (22.09%), Arts.Administration (21.93%), Politics (21.88%), School.Principals (21.75%), Social.Organizational.Psychology (21.19), Private.School.Leadership (20.48%), Cooperation.and.Conflict.Resolution (20.42%), and Leadership (20.06%) make up the highest percent variance within PC1. This simply means that these programs contribute their respective amounts to PC1. These programs are all related to leadership, organization, and administration concepts.

PC2: Clinical.Psychology (25.31%), Neuroscience (25.29%), Kinesiology (25.15), Physiology (24.43%), Psychology (22.37%), Health.Education (22.13%), Behavioral.Analysis (21.26%), Nursing (21.21%), Physical.Education (21.08%), and Counseling.Psychology (19.57%) are all within PC2 with relatively high percentages of variance making up most of PC2. These programs are all related to health including mental health or education of the subjects.

PC3: Design.and.Development.of.Digital.Games (31.52%), Cognitive.Science (31.04%), Mathematics (27.94%), Learning.Analytics (27.93), Education.Technology (26.90%), Creative.Technologies (26.10%), Instructional.Technology.and.Media (25.66%), and  Measurement.Evaluation.and.Statistics (24.67) are all within PC3 with relatively high percentages of variance making up most of PC3. These programs are all related to technology, data science, and statistical measures.

PC4: Linguistics (34.79), English.Education (34.07), Teaching.English (27.46), and Literacy (24.96) all are relatively high percentages of variance making up PC4 and all relate to learning language and reading. As you read down past Literacy and look at the other programs with high variance within PC4, they also are in some way related to learning.

PC5: History (32.73%) makes up the most of PC5 with the next closest percentage being Music (24.55%), which at least based on the program is not related to the subject of History and therefore PC5 is made up mostly of a program unrelated to the other classes within the component. The gap between them is large.

PC6: Science.Education (35.53%) and Higher.and.Postsecondary.Education (32.91%) make up the majority of PC6 variance, and are both related to education.